simodels: A package for Spatial Interaction Modelling

### 2024-09-04 paper/paper.qmd
Describe Leeds commute example

---
title: "simodels: A package for Spatial Interaction Modelling"
bibliography: bibliography.bib
execute: 
  messages: false
  warning: false 
  echo: false
---

@@

```{r}
od = simodels::si_od_census
zones = simodels::si_zones
nrow(od)
sum(od$all)
sum(zones$all)
```

>>>>

---
title: "simodels: A package for Spatial Interaction Modelling"
bibliography: bibliography.bib
number-sections: true
execute: 
  messages: false
  warning: false 
  echo: false
---

@@

The package comes with an example OD dataset from Leeds, UK, which can be used to demonstrate the SIMs.
Let's model travel to work in a major UK city.
We'll first load the inputs, representing administrative zones in Leeds (`si_zones`) and the OD data (`si_od_census`), which we will rename for clarity:

```{r}
#| echo: true
od = simodels::si_od_census
zones = simodels::si_zones
nrow(od)
sum(od$all)
sum(zones$all)
names(od)
```

As shown in the code and outputs above, the input dataset represents the commutes made by `r sum(od$all) |> format(big.mark = ",")` people in Leeds.
We have the breakdown of trips by mode, with `r sum(od$car_driver) |> format(big.mark = ",")` people driving to work, `r sum(od$car_passenger) |> format(big.mark = ",")` people car sharing, `r sum(od$bus) |> format(big.mark = ",")`, `r sum(od$bicycle) |> format(big.mark = ",")` cycling, and `r sum(od$foot) |> format(big.mark = ",")` people walking.
A hypothetical research question to inform policy with reference to this dataset is: "How would travel demand change if transport costs increase?".
Before answering this question, we need to develop a baseline model to predict the travel between origins and destinations as a function of distance (and other factors).
<!-- (Note: key other factors to test are 'network distance' representing intervening opportunities, and 'population' representing the number of potential travellers). -->
